---
title: "ppatil8_Homework3"
output:
  pdf_document: default
  html_document: default
date: "2024-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 4

```{r}
library(TSA)
```
#Time Plot of an MA(1) Process

```{r}
data(ma1.2.s)
data(ma1.2.s); plot(ma1.2.s,ylab=expression(Y[t]),type='o')
```

```{r}
plot(y=ma1.2.s,x=zlag(ma1.2.s),ylab=expression(Y[t]),
       xlab=expression(Y[t-1]),type='p')
```


```{r}
plot(y=ma1.2.s,x=zlag(ma1.2.s,2),ylab=expression(Y[t]),
     xlab=expression(Y[t-2]),type='p')
```

```{r}
data(ma1.1.s)
plot(ma1.1.s,ylab=expression(Y[t]),type='o')
```


```{r}
plot(y=ma1.1.s,x=zlag(ma1.1.s),ylab=expression(Y[t]),
       xlab=expression(Y[t-1]),type='p')
```

```{r}
plot(y=ma1.1.s,x=zlag(ma1.1.s,2),ylab=expression(Y[t]),
     xlab=expression(Y[t-2]),type='p')
```

#Time Plot of an MA(2) Process with
```{r}
data(ma2.s); plot(ma2.s,ylab=expression(Y[t]),type='o')
```


```{r}
plot(y=ma2.s,x=zlag(ma2.s),ylab=expression(Y[t]),
       xlab=expression(Y[t-1]),type='p')
```

```{r}
plot(y=ma2.s,x=zlag(ma2.s,2),ylab=expression(Y[t]),
     xlab=expression(Y[t-2]),type='p')
```


```{r}
plot(y=ma2.s,x=zlag(ma2.s,3),ylab=expression(Y[t]),
     xlab=expression(Y[t-3]),type='p')
```
#Time Plot of an AR(1) Series

```{r}
data(ar1.s); plot(ar1.s,ylab=expression(Y[t]),type='o')
```


```{r}
plot(y=ar1.s,x=zlag(ar1.s),ylab=expression(Y[t]),
       xlab=expression(Y[t-1]),type='p')
```

```{r}
plot(y=ar1.s,x=zlag(ar1.s,2),ylab=expression(Y[t]),
     xlab=expression(Y[t-2]),type='p')
```


```{r}
plot(y=ar1.s,x=zlag(ar1.s,3),ylab=expression(Y[t]),
     xlab=expression(Y[t-3]),type='p')
```
#Time Plot of an AR(2) Series

```{r}
data(ar2.s); 
plot(ar2.s,ylab=expression(Y[t]),type='o')
```

# Chapter 5

#Monthly Price of Oil: January 1986–January 2006

```{r}
data(oil.price)
plot(oil.price, ylab='Price per Barrel',type='l')
```

#An Explosive “AR(1)” Series

```{r}
data(explode.s)
plot(explode.s,ylab=expression(Y[t]),type='o')
```

#The Difference Series of the Logs of the Oil Price Time

```{r}
plot(diff(log(oil.price)),ylab='Change in Log(Price)',type='l')
```


```{r}
data(ima22.s)
plot(ima22.s,ylab='IMA(2,2) Simulation',type='o')
```

#First Difference of the Simulated IMA(2,2) Series

```{r}
plot(diff(ima22.s),ylab='First Difference',type='o')
```

#Second Difference of the Simulated IMA(2,2) Series

```{r}
plot(diff(ima22.s,difference=2),ylab='Differenced
Twice',type='o')
```

#U.S. Electricity Generated by Month

```{r}
data(electricity); plot(electricity)
```

#Time Series Plot of Logarithms of Electricity Values

```{r}
plot(log(electricity),ylab='Log(electricity)')
```

#Difference of Logarithms for Electricity Time Series

```{r}
plot(diff(log(electricity)),
     ylab='Difference of Log(electricity)')
```

#Log-likelihood versus Lambda

```{r}
BoxCox.ar(electricity)
```
