---
title: "ppatil8_Homework1"
output: pdf_document
date: "2024-02-09"
---

#12.5.1
```{r}
states <- row.names(USArrests)
states
names(USArrests)
apply(USArrests,2,mean)
apply(USArrests,2,var)
pr.out <- prcomp(USArrests,scale = TRUE)
names(pr.out)
pr.out$center
pr.out$scale
pr.out$rotation
dim( pr.out$x)
biplot(pr.out, scale = 0)

pr.out$rotation = -pr.out$rotation
pr.out$x = -pr.out$x
biplot( pr.out , scale = 0)
pr.out$sdev
pr.var <- pr.out$sdev^2
pr.var

pve <- pr.var/sum(pr.var)
pve

par(mfrow = c (1 , 2) )
plot(pve ,xlab = "Principal Component" ,
         ylab = "Proportion of Variance Explained" , ylim = c(0 , 1) ,
         type = "b" )
plot(cumsum(pve) , xlab = "Principal Component" ,
         ylab = "Cumulative Proportion of Variance Explained" ,
         ylim = c(0 , 1) ,type = "b" )
     
a <- c(1 , 2 , 8 , -3)
cumsum(a)
```

#12.5.3

```{r}
set.seed (2)
x <- matrix ( rnorm (50 * 2) , ncol = 2)
x [1:25 , 1] <- x [1:25 , 1] + 3
x [1:25 , 2] <- x [1:25 , 2] - 4
km.out <- kmeans(x , 2 , nstart = 20)
km.out $ cluster
par( mfrow = c (1 , 2))
plot(x , col = ( km.out $ cluster + 1) ,
      main = "K - Means Clustering Results with K = 2",
      xlab = " ",ylab = " " , pch = 20 , cex = 2)

set.seed(4)
km.out <- kmeans(x, 3, nstart = 20)
km.out
plot(x, col = (km.out$cluster + 1), main = "K-Means Clustering Results with K = 3", xlab = "", ylab = "", pch = 20, cex = 2)

set.seed(4)
km.out <- kmeans(x, 3, nstart = 1)
km.out$tot.withinss
km.out <- kmeans(x, 3, nstart = 20)
km.out$tot.withinss

#Hierarchical Clustering
hc.complete <- hclust(dist(x), method = "complete")
hc.average <- hclust(dist(x), method = "average")
hc.single <- hclust(dist(x), method = "single")
par(mfrow = c(1, 3))
plot(hc.complete, main = "Complete Linkage", xlab = "", sub = "", cex = .9)
plot(hc.average, main = "Average Linkage", xlab = "", sub = "", cex = .9)
plot(hc.single, main = "Single Linkage", xlab = "", sub = "", cex = .9)

cutree(hc.complete, 2)
cutree(hc.average, 2)
cutree(hc.single, 2)

cutree(hc.single, 4)

xsc <- scale(x)
plot(hclust(dist(xsc), method = "complete"), main = "Hierarchical Clustering with Scaled Features")

x <- matrix(rnorm(30 * 3), ncol = 3)
dd <- as.dist(1 - cor(t(x)))
plot(hclust(dd, ,method = "complete"), main = "Complete Linkage with Correation-Based Distance", xlab = "", sub = "")
```

#Exercise 12.5.3
#Q.8
#a
```{r}
pr.out <- prcomp(USArrests, scale = TRUE)
pr.var <- pr.out$sdev^2
pve <- pr.var / sum(pr.var)
sum(pr.var)
pve
```

#b
```{r}
loadings <- pr.out$rotation
USArrests2 <- scale(USArrests)
sumvar <- sum(apply(as.matrix(USArrests2)^2, 2, sum))
apply((as.matrix(USArrests2) %*% loadings)^2, 2, sum) / sumvar
```

#Q.9
#a
```{r}
set.seed(40)
hc.complete <- hclust(dist(USArrests), method = "complete")
plot(hc.complete)
```

#b
```{r}
cutree(hc.complete, 3)

```

#c
```{r}
sd.data <- scale(USArrests)
hc.complete.sd <- hclust(dist(sd.data), method = "complete")
plot(hc.complete.sd)
```

#d
```{r}
cutree(hc.complete.sd, 3)
table(cutree(hc.complete, 3), cutree(hc.complete.sd, 3))

```

#Scaling leads to diverse clusters, and the decision regarding whether to scale or not hinges on the specific data being analyzed.
#In part 9b and 9d, it can be seen how different clusters are chosed based on scaling
#Here for USArrests dataset, the variables Murder,Assault, UrbanPop and Rape are not on the same unit and for accurate results, it is necessary for them to be on the same scale hence in this case scaling is necessary