---
title: "ppatil8_Homework2"
output:
  pdf_document: default
  html_document: default
date: "2024-02-22"
---


#######Chapters 1-3#######

# Chapter1
#Time Series Plot of Los Angeles Annual Rainfall
```{r}
library(TSA)
data(larain)
plot(larain,ylab='Inches',xlab='Year',type='o')
```

#Scatterplot of LA Rainfall versus Last Year’s LA Rainfall

```{r}
plot(y=larain,x=zlag(larain),ylab='Inches',
     xlab='Previous Year Inches')
```

#Time Series Plot of Color Property from a Chemical Process

```{r}
data(color)
plot(color,ylab='Color Property',xlab='Batch',type='o')
```

#Scatterplot of Color Value versus Previous Color Value

```{r}
plot(y=color,x=zlag(color),ylab='Color Property',
     xlab='Previous Batch Color Property')
```

#Abundance of Canadian Hare

```{r}
data(hare); 
plot(hare,ylab='Abundance',xlab='Year',type='o')
```

#Hare Abundance versus Previous Year’s Hare Abundance

```{r}
plot(y=hare,x=zlag(hare),ylab='Abundance',
     xlab='Previous Year Abundance')
```

#Average Monthly Temperatures, Dubuque, Iowa

```{r}
data(tempdub); plot(tempdub,ylab='Temperature',type='o')
```
#Monthly Oil Filter Sales

```{r}
data(oilfilters); plot(oilfilters,type='o',ylab='Sales')
```

#Monthly Oil Filter Sales with Special Plotting Symbols

```{r}
plot(oilfilters,type='l',ylab='Sales')
points(y=oilfilters,x=time(oilfilters),
       pch=as.vector(season(oilfilters)))
```

# Chapter2
#Time Series Plot of a Random Walk

```{r}
data(rwalk)
plot(rwalk,type='o',ylab='Random Walk')
```

# Chapter3
# Least Squares Regression Estimates for Linear Time Trend

```{r}
data(rwalk)
model1=lm(rwalk~time(rwalk))
summary(model1)
```

#Random Walk with Linear Time Trend

```{r}
plot(rwalk,type='o',ylab='y')
abline(model1)
```

#Regression Results for the Seasonal Means Model

```{r}
data(tempdub)
month.=season(tempdub)
model2=lm(tempdub~month.-1)
summary(model2)
```

#Results for Seasonal Means Model with an Intercept

```{r}
model3=lm(tempdub~month.) 
summary(model3)
```

#Cosine Trend Model for Temperature Series

```{r}
har.=harmonic(tempdub,1)
model4=lm(tempdub~har.)
summary(model4)
```

#Cosine Trend for the Temperature Series

```{r}
plot(ts(fitted(model4),freq=12,start=c(1964,1)),
     ylab='Temperature',type='l',
     ylim=range(c(fitted(model4),tempdub)));
points(tempdub)
```

#Regression Output for Linear Trend Fit of Random Walk

```{r}
model1=lm(rwalk~time(rwalk))
summary(model1)
```

#Residuals versus Time for Temperature Seasonal Means

```{r}
plot(y=rstudent(model3),x=as.vector(time(tempdub)),
     xlab='Time',ylab='Standardized Residuals',type='o')
```

#Residuals versus Time with Seasonal Plotting Symbols

```{r}
plot(y=rstudent(model3),x=as.vector(time(tempdub)),
     xlab='Time',ylab='Standardized Residuals',type='l')
points(y=rstudent(model3),x=as.vector(time(tempdub)),
       pch=as.vector(season(tempdub)))
```

#Standardized Residuals versus Fitted Values for the Temperature Seasonal Means Model

```{r}
plot(y=rstudent(model3),x=as.vector(fitted(model3)),
     xlab='Fitted Trend Values', ylab='Standardized Residuals',type='n')
points(y=rstudent(model3),x=as.vector(fitted(model3)),
       pch=as.vector(season(tempdub)))
```

#Histogram of Standardized Residuals from Seasonal Means Model

```{r}
hist(rstudent(model3),xlab='Standardized Residuals')
```

#Q-Q Plot: Standardized Residuals of Seasonal Means Model

```{r}
qqnorm(rstudent(model3))
```

#Sample Autocorrelation of Residuals of Seasonal Means Model

```{r}
acf(rstudent(model3))
```

#Residuals from Straight Line Fit of the Random Walk

```{r}
plot(y=rstudent(model1),x=as.vector(time(rwalk)),
     ylab='Standardized Residuals',xlab='Time',type='o')
```

#Residuals versus Fitted Values from Straight Line Fit

```{r}
plot(y=rstudent(model1),x=fitted(model1),
     ylab='Standardized Residuals',xlab='Fitted Trend Line Values',
     type='p')
```

#Sample Autocorrelation of Residuals from Straight Line Model

```{r}
acf(rstudent(model1))
```

#Quantile-Quantile Plot of Los Angeles Annual Rainfall Series

```{r}
qqnorm(larain); qqline(larain)
```

#######Exercise#######

#Ex. 1.1

```{r}
library(TSA)
library(latticeExtra)
data(larain, package = "TSA")
xyplot(larain, ylab = "Inches", xlab = "Year", type = "o")
```

#Ex. 1.2

```{r}
data(color)
xyplot(color, ylab = "Color property", xlab = "Batch", type = "o")
```

#Ex. 1.3

```{r}
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
xyplot(as.ts(rnorm(48)))
```

#No pattern detected


